---
title: "COVID-19"
output: 
  flexdashboard::flex_dashboard:
    theme: cosmo
    orientation: rows
    vertical_layout: scroll
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(ggplot2)
library(tidyr)
library(dplyr)
library(leaflet)
library(plotly)
library(stringr)
library(RCurl)
library(tidyverse)
library(data.table)
library(maps)
library(geojson)
library(geojsonio)
library(geojsonlint)

#Import data:
world <- read.csv("COVID19_March_16_case_count.csv")

```
World Overview
=====================================  

Row
-----------------------------------------------
### Confirmed cases

```{r}

renderValueBox({
  confirmed_cases <- sum(world$Confirmed)
  valueBox(prettyNum(confirmed_cases, big.mark = ","),
           color = "maroon")
})
```

### Deaths 
```{r}
renderValueBox({
  deaths <- sum(world$Deaths)
  valueBox(prettyNum(deaths, big.mark = ","), 
           color = "black")
})
```

### Recovered 
```{r}
renderValueBox({
  recovered <- sum(world$Recovered)
  valueBox(prettyNum(recovered, big.mark = ","),
           color = "success")
})
```

Row {data-height=650}
--------------------------------------------
### Distribution of culmulative cases worldwide
    
```{r}

leaflet() %>% 
  addTiles() %>% 
  addCircleMarkers(world$Longitude, world$Latitude, 
                   radius = sqrt(world$Confirmed)/8,
                   fillColor = 'red',
                  color = "red",
                   weight = 3,
                   fill = T,
                   fillOpacity = 0.5,
                   opacity = 0.5,
                   popup = paste(world$Province.State, sep = " ", world$Country.Region, 
                                 world$Confirmed, "confirmed cases"))

```

Row {data-height = 50}
--------------------------------------------
### Search by country or region

```{r}

cases_df = world %>%
    select(Country.Region, Confirmed, Deaths, Recovered) 

cases_df1 = aggregate(x = cases_df[, 2:4], by = list(cases_df$Country.Region), FUN = sum)
colnames(cases_df1) = c("Country or Region", "Confirmed", "Deaths", "Recovered")
cases_df1 = cases_df1 %>% arrange(desc(Confirmed))
DT::renderDataTable({
DT::datatable(cases_df1,  rownames = TRUE, options = list(scrollY = '300px',pageLength = 5))
})
```

### Total number of cases worldwide over time
```{r}
time_series_confirmed = read.csv("time_series_19-covid-Confirmed.csv")
time_series_death = read.csv("time_series_19-covid-Deaths.csv")
time_series_recovered = read.csv("time_series_19-covid-Recovered.csv")

time_series_confirmed1 = aggregate(x = time_series_confirmed[, 5:59], by = list(time_series_confirmed$Country.Region), FUN = sum)
time_series_death1 = aggregate(x = time_series_death[, 5:59], by = list(time_series_death$Country.Region), FUN = sum)
time_series_recovered1 = aggregate(x = time_series_recovered[, 5:59], by = list(time_series_recovered$Country.Region), FUN = sum)

# calculate global trend on cases
dates = seq(as.Date("2020/1/22"), as.Date("2020/3/16"), "days")
confirmed_cases = colSums(time_series_confirmed1[,-1])
death = colSums((time_series_death1[,-1]))
recovered = colSums(time_series_recovered1[,-1])

global_trend = data.frame(Date = dates, confirmed_cases = confirmed_cases, death = death, recovered = recovered)
colnames(global_trend) = c("Date", "Confirmed cases", "Deaths", "Recovered")

p = ggplot() + 
  geom_line(data = global_trend, aes(x = Date, y = `Confirmed cases`), color = "maroon", size = 1.5) +
  geom_line(data = global_trend, aes(x = Date, y = Deaths), color = "black", size = 1.5) +
  geom_line(data = global_trend, aes(x = Date, y = Recovered), color = "green", size = 1.5) +
  xlab('Date') +
  ylab('Number of cases') +
  theme_minimal() +
  scale_fill_discrete(name="Total cases over time",
                     labels=c("Confirmed cases", "Deaths", "Recovered"))

ggplotly(p)
```

Row
-----------------------------------------------
### **Footnotes:** 
Created by **Vy Tran**. Case information was updated on **March 16, 2020**. Codes and data are available on [Github](https://github.com/vy-p-tran/COVID19_Vietnam).<br />
Data was obtained from the [Johns Hopkins Center for Systems Science and Engineering](https://github.com/CSSEGISandData/COVID-19). Their popular dashboard can be found [here](https://www.arcgis.com/apps/opsdashboard/index.html#/bda7594740fd40299423467b48e9ecf6).


Vietnam 
===================================== 


Row
-----------------------------------------------
### Confirmed cases

```{r}
Vietnam = read.csv("Vietnam_cases_March_16.csv")
renderValueBox({
  confirmed_cases <- sum(Vietnam$Confirmed)
  valueBox(prettyNum(confirmed_cases, big.mark = ","),
           color = "maroon")
})
```

### Deaths 
```{r}
renderValueBox({
  deaths <- sum(Vietnam$Death)
  valueBox(prettyNum(deaths, big.mark = ","), 
           color = "black")
})
```

### Recovered 
```{r}
renderValueBox({
  recovered <- sum(Vietnam$Recovered)
  valueBox(prettyNum(recovered, big.mark = ","),
           color = "success")
})
```

Row {data-height=650}
-------------------------------------

### Distribution of cumulative cases in Vietnam
```{r}

# Create interactive labels for leaflet graph
labels <- sprintf(
  "<strong>%s</strong><br/>%s Confirmed <br/>%s Deaths<br/>%s Recovered",
  Vietnam$Province.City, Vietnam$Confirmed, Vietnam$Death, Vietnam$Recovered) %>% 
  lapply(htmltools::HTML)


# create a leaflet map on which we will build
leaflet() %>% 
  setView( lng = 106.7,
            lat = 15,
            zoom = 5.49) %>%
  addTiles() %>% 
  addCircleMarkers(Vietnam$lng, 
                   Vietnam$lat, 
                   radius = Vietnam$Confirmed,
                   fillColor = 'red',
                   color = 'red',
                   weight = 3,
                   fill = T,
                   fillOpacity = 0.6,
                   opacity = 0.6, label = labels)

```

Row
-----------------------------------------------
### Cases in Vietnam by province and city

```{r}
Vietnam_cases = Vietnam %>%
    select(Province.City, Confirmed, Death, Recovered) 
colnames(Vietnam_cases) = c("Province or City", "Confirmed", "Death", "Recovered")
Vietnam_cases = Vietnam_cases%>% arrange(desc(Confirmed))
DT::renderDataTable({
DT::datatable(Vietnam_cases,  rownames = TRUE, options = list(scrollY = '300px',pageLength = 13))
})
```

### Total number of cases in Vietnam over time

```{r}
vn_confirmed1 = filter(time_series_confirmed1, Group.1 == "Vietnam")
vn_death1 = filter(time_series_death1, Group.1 == "Vietnam")
vn_recovered1 = filter(time_series_recovered1, Group.1 == "Vietnam")

# calculate Vietnam trend on cases
dates = seq(as.Date("2020/1/22"), as.Date("2020/3/16"), "days")
confirmed_cases = colSums(vn_confirmed1[,-1])
death = colSums((vn_death1[,-1]))
recovered = colSums(vn_recovered1[,-1])
vn_trend = data.frame(Date = dates, confirmed_cases = confirmed_cases, death = death, recovered = recovered)
colnames(vn_trend) = c("Date", "Confirmed cases", "Deaths", "Recovered")

p = ggplot() + 
  geom_line(data = vn_trend, aes(x = Date, y = `Confirmed cases`), color = "maroon", size = 1.5) +
  geom_line(data = vn_trend, aes(x = Date, y = Deaths), color = "black", size = 1.5) +
  geom_line(data = vn_trend, aes(x = Date, y = Recovered), color = "green", size = 1.5) +
  xlab('Date') +
  ylab('Number of cases') +
  theme_minimal() +
  scale_fill_discrete(name="Total cases over time",
                     labels=c("Confirmed cases", "Deaths", "Recovered"))

ggplotly(p)
```

Row
-----------------------------------------------
### **Footnotes:**
Information was updated on **March 16, 2020**. Data was collected from [Vietnam Minister of Health](https://ncov.moh.gov.vn/).


United States
==================================================

Row 
-----------------------------------------------
### Confirmed cases

```{r}
# Data pre-processing for US cases:
COVID19_case_count = read.csv("COVID19_March_16_case_count.csv")
US1 = filter(COVID19_case_count, Country.Region == "US") 
setnames(US1, "Province.State", "States")
US2 = select(US1, States, Confirmed, Deaths, Recovered, Latitude, Longitude)
US2 = US2[order(US2$States),]
US2 = US2[-c(9,13, 14, 51),]

renderValueBox({
  confirmed_cases <- sum(US2$Confirmed)
  valueBox(prettyNum(confirmed_cases, big.mark = ","),
           color = "maroon")
})
```

### Deaths 
```{r}
renderValueBox({
  deaths <- sum(US2$Deaths)
  valueBox(prettyNum(deaths, big.mark = ","), 
           color = "black")
})
```

### Recovered 
```{r}
renderValueBox({
  recovered <- sum(US2$Recovered)
  valueBox(prettyNum(recovered, big.mark = ","),
           color = "success")
})
```

Row {data-height = 650}
-------------------------------------------------
### Distribution of cumulative cases in the US: 
```{r}
### Distribution of cases in the US by state

# Get the US map data:
url <- "http://leafletjs.com/examples/choropleth/us-states.js"
# Read it as text file:
doc <- readLines(url)
# remove the javascript assignment at the front 
doc2 <- gsub("var statesData = ", "", doc)
# write out as a temp file and read
write(doc2, file = "tempgeo.json")
states <- geojson_read("tempgeo.json", what = "sp")
# Match states in json file with cases in US2:
a = as.data.frame(states$name)
colnames(a) = "States"
b = as.data.frame(select(US2, States, Confirmed, Deaths, Recovered))
c = left_join(a, b, by  = "States")
states$density = c$Confirmed
# Create interactive labels for leaflet graph
labels <- sprintf(
  "<strong>%s</strong><br/>%s Confirmed <br/>%s Deaths<br/>%s Recovered",
  c$States, c$Confirmed, c$Deaths, c$Recovered) %>% 
  lapply(htmltools::HTML)
bins = c(0, 10, 20, 30, 40, 60, 80, 100, 200,400, 600, 800)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)
# create a leaflet map on which we will build
leaf <- leaflet(states) %>%
    setView(-96, 37.8, 4) %>%
    addTiles() %>%
addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
addPolygons(fillColor = ~pal(states$density), weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,label = labels) %>%
addLegend(pal = pal, values = ~states$density, opacity = 0.7, title = NULL,
  position = "bottomright")
leaf
```


Row {data-height=350}
----------------------------------------
### States with available COVID-19 testing
```{r}
COVID19_US_Testing = read.csv("COVID19_US_Testing.csv")
states2 = states
states2$density = COVID19_US_Testing$Have_test
pal2 <- colorFactor(palette = c("lightgreen", "red"), COVID19_US_Testing$Have_test)
# create a leaflet map on which we will build
labels2 <- sprintf(
  "<strong>%s<br/></strong>%s",
  states2$name, states2$density) %>% 
  lapply(htmltools::HTML)
leaf2 <- leaflet(states2) %>%
    setView(-96, 37.8, 4) %>%
    addTiles() %>%
addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
addPolygons(stroke = FALSE, smoothFactor = 0.2, color = pal2(COVID19_US_Testing$Have_test), weight = 2,
  opacity = 1,
  dashArray = "3",
  fillOpacity = 0.7,label = labels2) %>%
addLegend(pal = pal2, values = ~COVID19_US_Testing$Have_test, opacity = 0.7, title = NULL,
  position = "bottomright")
leaf2
```

### Cases of COVID-19 by sources of exposure in the US
```{r }
#Data was obtained from the CDC on 03/15/2020:
exposure = data.frame("Sources of Exposure" = c("Travel-related", "Close contact", "Under investigation"), Value = c(205, 214, 3068), Percent = c(5.88, 6.14, 87.99))
exposure = exposure[order(exposure$Sources.of.Exposure),]
pie = ggplot(exposure, aes(x = "", y= Percent, fill = Sources.of.Exposure)) + 
  geom_bar(width = 1, stat = 'identity') +
  coord_polar("y") +
  scale_fill_brewer(name="Sources of exposure for confirmed cases", labels = c("Travel-related 205 cases", "Close contact 214 cases","Under investigation 3068 cases"), palette = 'Oranges') + 
  theme(axis.text.x=element_blank(),  legend.text=element_text(size=10)) + theme_void() +
  geom_text(aes(label = paste0((Percent), "%")),
              position = position_stack(vjust = 0.5), size = 3) 
pie
```


Row 
-------------------------------------
### **Footnotes:**
Data was last updated **March 16, 2020**.<br />
According to the [CDC](https://www.cdc.gov/coronavirus/2019-ncov/cases-updates/testing-in-us.html), as of the morning of March 17th, 84 state and local public health laboratories in 50 states, the District of Columbia, and Guam have successfully verified COVID-19 diagnostic tests and are offering testing.

